---
title: "Exploratory Factor Analysis"
author: "Heru Wiryanto"
date: "August 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# WHAT'S FACTOR ANALYSIS
Factor analysis can best be understood as a latent variable modeling paradigm in which a set of observed variables are the indicators of a latent variable. In this schema, the latent variable (e.g. intelligence) is of primary interest, but cannot be directly observed. However, it is theorized that the latent variable has a direct influence on each of the observed indicators (e.g. items on a scale, subscales in a battery of measures), so that they can in turn be used to gain insights into the latent variable. This idea is at the core of educational and psychological measurement of abilities.
library(foreign)

#EXPLORATORY FACTOR ANALYSIS
To begin, let us consider an example in which a researcher has collected data on
achievement goal orientation using the 12-item Likert achievement goal scale.  Each item has seven options ranging from “not at all like me” to “very true of me”. 
##ITEMS
The items appear below.
 - AGS1 = My goal is to completely master the material presented in my classes. (MAP)
 
 - AGS2 = I want to avoid learning less than it is possible to learn. (MAV)
 
 - AGS3 = It is important for me to do better than other students. (PAP)
 
 - AGS4 = I want to avoid performing poorly compared to others. (PAV)
 
 - AGS5 = I want to learn as much as possible. (MAP)
 
 - AGS6 = It is important for me to avoid an incomplete understanding of the course material. (MAV)
 
 - AGS7 = It is important for me to understand the content of my courses as thoroughly as possible. (MAP)
 
 - AGS8 = My goal is to avoid performing worse than other students. (PAV)
 
 - AGS9 = I want to do well compared to other students. (PAP)
 
 - AGS10 = It is important for me to avoid doing poorly compared to other students. (PAV)
 
 - AGS11 = My goal is to perform better than the other students. (PAP)
 
 - AGS12 = My goal is to avoid learning less than I possibly could. (MAV)

The researcher would like to investigate the latent structure of achievement goal orientation, using the responses to these 12 items from 430 college students. The theory underlying the AGS states that there exist four distinct latent traits: mastery approach (MAP), mastery avoidant (MAV), performance approach (PAP),and performance avoidant (PAV)

##FITTING EFA MODEL WITH FACTANAL 
```{r, echo=FALSE}
data<-read.csv("MyData.csv", header=TRUE, sep=,)
Mydata <- na.omit(data)
performance.data<-Mydata
attach(performance.data)
achievement.goal.efa.results4<-factanal(~ags1+ags2+ags3+ags4+ags5+ags6+ags7+ags8+ags9+ags10+ags11+ags12,factors=4, rotation="promax")
achievement.goal.efa.results4
```
for example, approximately 48.7% of the variance in item ags1 is not associated with the four
retained factors.
```{r, echo=FALSE}
1-achievement.goal.efa.results4$uniquenesses
```
Thus, approximately 51% of the variance in ags1 is associated with the four factors, compared to 99.5% of the factor accounted for variance in ags8.

Finally, the chi-square goodness of fit test is the last portion of the output from factanal.
For this model, the chi-square was 77.4 with 24 degrees of freedom, and a p-value of 0.000132. This value is well below our α of 0.05, **leading us to reject the null hypothesis that the model adequately fits the data.** 

## Including Plots

You can also embed plots, for example:
```{r, echo=FALSE}
library(psych)
library(nFactors)
ev <- eigen(cor(performance.data)) # get eigenvalues
ap <- parallel(subject=nrow(performance.data),var=ncol(performance.data),rep=100,cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```

##FACTANAL DENGAN 3 FACTOR DAN PROMAX
```{r, echo=FALSE}
achievement.goal.efa.results3<-factanal(~ags1+ags2+ags3+ags4+ags5+ags6+ags7+ags8+ags9+ags10+ags11+ags12, factors=3, rotation="promax")
achievement.goal.efa.results3
```

##FACTANAL DENGAN 2 FACTOR DAN PROMAX
```{r, echo=FALSE}
achievement.goal.efa.results2<-factanal(~ags1+ags2+ags3+ags4+ags5+ags6+ags7+ags8+ags9+ags10+ags11+ags12, factors=2, rotation="promax")
achievement.goal.efa.results2
```

##FITTING EFA DENGAN FA DENGAN 4 FACTOR
```{r, echo=FALSE}
achievement.goal.efa.results4.paf<-fa(performance.data, nfactors=4, residuals=TRUE, rotate="promax", SMC=TRUE, fm="pa")
achievement.goal.efa.results4.paf
```

##FITTING EFA DENGAN FA DENGAN 3 FACTOR
```{r, echo=FALSE}
achievement.goal.efa.results3.paf<-fa(performance.data, nfactors=3, residuals=TRUE, rotate="promax", SMC=TRUE, fm="pa")
achievement.goal.efa.results3.paf
```

##FITTING EFA DENGAN FA DENGAN 2 FACTOR
```{r, echo=FALSE}
achievement.goal.efa.results2.paf<-fa(performance.data, nfactors=2, residuals=TRUE, rotate="promax", SMC=TRUE, fm="pa")
achievement.goal.efa.results2.paf
```

##FITTING EFA DENGAN PCA
```{r, echo=FALSE}
achievement.goal.pca<-princomp(~ags1+ags2+ags3+ags4+ags5+ags6+ags7+ags8+ags9+ags10+ags11+ags12)
achievement.goal.pca$loadings
```

